---
title: "Week 1 - Basic Statistics"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
help(co2)
```

Wrap in extra () to print.

Read: co2 ~ time(co2) as 'co2 on time(co2)'

```{r}
(co2.linear.model = lm(co2 ~ time(co2)))
```
```{r}
summary(co2.linear.model)
```

```{r}
sum(co2.residuals^2)
```
```{r}
sum(resid(co2.linear.model)^2)

```

```{r}
{plot(co2, main='Atmospheric CO2 Concentration with Fitted Line')
abline(co2.linear.model)
}
```

## Histogram of Residuals

```{r}
co2.residuals = resid(co2.linear.model)
hist(co2.residuals, main="Histogram of Residuals")
```

Don't use a histogram when you only have a few points.

## qqnorm

```{r}
{qqnorm(co2.residuals)
qqline(co2.residuals)
}
```

We see departures from normality on both ends.


## Time Plot of Residuals
```{r}
plot(co2.residuals ~ time(co2))
```

# Faraway Quiz Example

Yes, they're normal
```{r}
library(faraway)
{plot(punting$Distance ~ punting$Hang); 
abline( lm(punting$Distance ~ punting$Hang) )
}
```

## Zoomed Residuals

```{r}
plot(co2.residuals ~ time(co2), xlim=c(1960,1963), main="Zoomed Residuals")
```

# Basic Statistics III

## Gosset Data

```{r}
help("sleep")
```

Second drug has an advantage over the first

```{r}
plot(extra ~ group, data = sleep, main="Gosset Data")
attach(sleep)
extra.1 = extra[group==1]
extra.2 = extra[group==2]
```

## Two-sided t-Test

```{r}
t.test(extra.1, extra.2, paired = TRUE, alternative = "two.sided")
```
df = 9, which is n-1

## Question

```{r}
bpm.before = c(64, 45, 51, 56, 76, 77, 48, 65, 88, 37, 49, 71, 63, 57, 43); 
bpm.after = c(82, 63, 64, 71, 91, 99, 66, 85, 100, 52, 59, 84, 73 ,75, 58 ); 
{qqnorm( bpm.after - bpm.before); 
qqline(bpm.after - bpm.before)
}
```

# Basic Statistics IV - Correlation

```{r}
help(trees) # Cherry Trees
```

## Pairs Plot
```{r}
pairs(trees, pch=21, bg=c("red"))
```

- Girth is a great predictor of volume
- Height is also a decent predictor of volume

## Covariance Matrix

```{r}
cov(trees)
```

Surprise: Covariance between height and volume is more (62.66)

Not consistent with picture, unless we think about units.

Covariance changes, depending on units
